---
title: "FAQ"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FAQ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ReporteRs)
library(magrittr)
if(!require(filelink))
  as_link <- function(file, txt, label){invisible()}

options("ReporteRs-fontsize" = 11)
```

## Installation

Install ReporteRs from cran with the following command:

```{r eval=FALSE}
install.packages("ReporteRs")
```


Install developement version of ReporteRs from Github with the following command:

```{r eval=FALSE}
devtools::install_github("davidgohel/ReporteRs")
```

### Dependancies

`ReporteRs` needs packages `rvg` and `gdtools`. 

**This requires a cpp coompiler.** If have no idea what it means :

* you're using R for Windows, install [Rtools](https://cran.r-project.org/bin/windows/Rtools)
* you're using R for Mac, install [Xcode](https://itunes.apple.com/us/app/xcode/id497799835)
* you're using R for Linux, you should manage on your own!

### Something screwed up

#### Could not find build tools

If you get an error during this step 
saying `Error: Could not find build tools necessary to build XXX`, you are probably installing the package from source (and rebuilding also rvg or gdtools), best advice is to read again that [part](#dependancies) :). 


#### rJava hell

Common errors are:

* JAVA_HOME cannot be determined from the Registry

```
Loading required package: rJava
library(rJava)
Error : .onLoad failed in loadNamespace() for ‘rJava’, details:
call: fun(libname, pkgname)
error: JAVA_HOME cannot be determined from the Registry
```

* unable to load shared object [...] rJava.dll

```
Error : .onLoad failed in loadNamespace() for 'rJava', details:
  call: inDL(x, as.logical(local), as.logical(now), ...)
  error: unable to load shared object '[...]/rJava/libs/x64/rJava.dll':
  LoadLibrary failure:  %1 is not a valid Win32 application.

Error: package/namespace load failed for 'rJava'
```

Most of the time, it comes from a bad java installation on your Windows machine:

* you are running R 64 bits but you have installed java 32 bits
* you are running R 32 bits but you have installed java 64 bits

Solution is to install the java version corresponding to your R version. I found that post very usefull:

https://www.r-statistics.com/2012/08/how-to-load-the-rjava-package-after-the-error-java_home-cannot-be-determined-from-the-registry/

I have a Windows 64 bits, I installed java 32 bits and java 64 bits. When I install ReporteRs, I don't meet any trouble. 

If you are on a Windows machine and (for any reason) you can only have one java version (32 or 64 bits), install the package with the following option :

```{r eval=FALSE}
install.packages("ReporteRs", INSTALL_opts = "--no-multiarch")
install_github("davidgohel/ReporteRs", args = "--no-multiarch")
```


Some solutions on Stackoverflow involve to set environment variable `JAVA_HOME`. I never had to use that but it seems a valid solution for some configurations. 

## Word: replace only a word in a paragraph

That is not a feature but this can be achieved by formating your text with `pot` objects. 

```{r}
library(ReporteRs)

# various text formatting properties we will use----
base_text_prop <- textProperties(font.family = "Arial")
bold_text_prop <- chprop( base_text_prop, 
                          font.weight = "bold", color = "red")
italic_text_prop <- chprop( base_text_prop, font.style = "italic")

# concatenate chunk of formatted text -----
your_pot <- pot("This is a pot ", format = base_text_prop) + 
  pot( "with italic text", format = italic_text_prop) + 
  pot( " and red bold", format = bold_text_prop)
  

# replace whole paragraph by the previous object -----
doc <- docx( title = "My example",
            template = file.path( system.file(package = "ReporteRs"),
                                  "templates/bookmark_example.docx") ) 

doc <- addParagraph(doc, your_pot, bookmark = "REVIEWER" )
target_file <- tempfile(fileext = ".docx") # file to produce 
writeDoc(doc, target_file)
```

```{r echo=FALSE}
as_link(file = target_file, txt = "It produces ", label = basename(target_file) )
```

## PowerPoint: offx, offy, width and height

You may want to arrange plot, tables or paragraphs freely on a slide - without mapping content to a layout defined in the slide master layouts. 

In this case, usage of some parameters is becoming mandatory. You will need to specify:

- `width` and `height` that are the shape dimensions 
- arguments `offx` and `offy` that are the shape positions (left and top positions)

## PowerPoint and par.properties

Usually there is no need for managing `par.properties` when adding content in a slide: i.e. graphics and tables cannot be associated with center or padding properties. Paragraphs can.

When formatting properties of paragraphs don't inherit from the template (**when adding a `pot` or a `set_of_paragraphs`**), `par.properties` can be used to specify alignment of paragraphs. 

```{r}
library(ReporteRs)

doc <- pptx( )
doc <- addSlide( doc, slide.layout = "Title and Content" )

doc <- addParagraph( par.properties = parCenter(),
    doc = doc, value = pot("this is a pot", textBold(color = "red")), 
    offx = 3, offy = 5, width = 2, height = 1 )

target_file <- tempfile(fileext = ".pptx") # file to produce 
writeDoc(doc, target_file)
```

```{r echo=FALSE}
as_link(file = target_file, txt = "It produces ", label = basename(target_file) )
```

