---
title: "Formatting properties"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formatting properties}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ReporteRs)
library( magrittr )
```

`ReporteRs` has functions to manage objects representing:

* text formatting properties
* paragraph formatting properties
* border formatting properties
* cell formatting properties.

With ReporteRs, some options can be used to reduce usage of some parameters. These are 
then used as default values for `textProperties` objects:

```{r}
getOption('ReporteRs-fontsize')
getOption('ReporteRs-default-font')
```

## Available formatting properties

### Text formatting properties

Create a `textProperties` object that describes text formatting properties. It
let you specify font name color, size, weight, etc.

`textProperties` objects can be used with function `pot` and when manipulating `FlexTable` objects.

#### Usage example

```{r}
my_text = pot('My tailor', textProperties(color='#1163A5', font.size = 10) ) +
  pot(' is ', textProperties(font.style = 'italic', color='gray', font.size = 12) ) +
  pot('rich', textProperties(font.weight='bold', color='#D63C3A', font.size = 10) )
my_text

MyFTable = vanilla.table( data = mtcars[1:5, ], add.rownames = TRUE )
MyFTable[ 1, 1, to = "header"] = my_text
MyFTable

```

### Paragraph formatting properties

Create a `parProperties` object that describes paragraph formatting properties.
It let you specify text alignment, paddings, the kind of numbering if paragraphs are a list, etc.

A `parProperties` object is used by several functions, ie. addPlot, addFlexTable, addParagraph.

#### Usage

```{r}
MyFTable = FlexTable( data = iris[1:5, ] )
MyFTable[ 1:2, 4:5] = parProperties( text.align = 'right', padding = 3 )
MyFTable
```


### Cell formatting properties

Create a `cellProperties` object that describes cell formatting properties.
It let you specify cell borders, background color, etc.

A cellProperties object is used with `FlexTable` objects.

#### Usage

```{r}
MyFTable = FlexTable( data = iris[1:5, ] )
MyFTable[ 1:2, 4:5] = cellProperties( background.color = 'wheat' )
MyFTable
```


### Border formatting properties

Create a `borderProperties` object that describes border formatting properties.

A borderProperties object is used with `FlexTable` objects and with `parProperties`
objects.

#### Usage

```{r}
MyFTable = FlexTable( data = iris[1:5, ] )
MyFTable[ 1, to = "header", side = "bottom"] = borderProperties(color="orange", style="solid", width=4)
MyFTable
```


## Change properties

To get modified copies of textProperties, parProperties, and cellProperties objects,
use the function `chprop`.

It makes the code less verbose and makes easier to specify formatting properties on FlexTable.

```{r}
base_text_prop = textProperties( font.size = 10, color = "#3D3234" )
base_par_prop = parProperties( text.align = "right" )
base_cell_prop = cellProperties( padding = 0 )

MyFTable <- vanilla.table( data = iris[1:5, ])
MyFTable[] <- base_par_prop
MyFTable[ , 5 ] <- chprop( base_par_prop, text.align = "center")
MyFTable[] <- base_cell_prop
MyFTable[ 3, ] <- chprop( base_cell_prop, background.color = "orange" )
MyFTable[] <- base_text_prop
MyFTable[ , 1] <- chprop( base_text_prop, font.weight = "bold" )
MyFTable
```

## Shorcut functions

The following functions are shortcuts for `textProperties` function calls:

* **textNormal**: return default textProperties object.
* **textBold**: return a textProperties object with bold font.
* **textBoldItalic**: return a textProperties object with bold italic font.
* **textItalic**: return a textProperties object with italic font.

The following functions are shortcuts for `parProperties` function calls:

* **parRight**: return a parProperties object with right text alignment.
* **parLeft**: return a parProperties object with left text alignment.
* **parJustify**: return a parProperties object with justified text alignment.
* **parCenter**: return a parProperties object with center text alignment.

```{r}
MyFTable = vanilla.table( data = iris[1:5, ] )
MyFTable[ to = "header" ] = textBold()
MyFTable[ , 5 ] = textItalic()
MyFTable[ , 1:4] = parRight()
MyFTable
```

The following functions are shortcuts for `borderProperties` function calls: **borderDashed**,
**borderDotted**, **borderSolid** and **borderNone**.

The following functions are shortcuts for `cellProperties` function calls: **cellBorderNone**, **cellBorderBottom**, **cellBorderTop** and **cellBorderTB**.


