---
title: "pot objects"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{pot objects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ReporteRs)
library( magrittr )
options("ReporteRs-fontsize" = 11)
```

A `pot` object (pieces of text or paragraph of text) is made 
of one or more chunk of text with different formatting properties. 

## Overview

ReporteRs provide a class of object to enable text formatting: `pot` objects. 

`pot` objects are created with function `pot`. These objects can be 
concatenated: an object can be added with another pot object or with a single 
text value (a character vector of length 1).

**A `pot` object is a set of text chunks, each chunk is 
associated with its own font properties.** A Footnote 
and an hyperlink can be associated to a `pot` object.

When `\\n` is used, a *soft return* is used.

These objects are to be used when adding paragraphs into a document or when working with FlexTable.

```{r}
pot('Hello World!')
pot('Hello') + ' ' + pot( 'World!', textItalic() )
pot( 'Cats', textProperties( color = 'orange' ) ) + ' and ' + 
  pot('dogs', textBold( color = '#E74C3C' ) )
```


## Usage

The function pot has two arguments:

* **value**: a text value or a value that has a format method returning character value.
* **format** (optional): formating properties (an  textProperties` object).

The following code show usage of these parameters:

```{r}
my_text_prop <- textProperties(color='#1163A5',
  font.size = 20, font.weight = 'bold', font.family = 'Courier New' )

my_text <- pot( value = 'My tailor', format = my_text_prop ) +
  ' is rich. ' +
  pot( value = 'Cats', format = chprop( my_text_prop, color = '#F0A91B') ) +
  ' and ' +
  pot( value = 'dogs', format = chprop( my_text_prop, color = '#D63C3A') )
my_text
```

## Examples

**add paragraphs**

```{r}
mydoc = docx( )
mydoc = addParagraph( mydoc, value = my_text )

```

It produces this [document](./files/pot_objects/word_pot_demo.docx).


**add into FlexTable**

```{r}

MyFTable = FlexTable( data = mtcars[1:5, ], add.rownames = TRUE )
MyFTable = addFooterRow( MyFTable, value = '',colspan = ncol(mtcars) + 1)
MyFTable[1,1, to = 'footer'] = my_text

MyFTable
```

## Hyperlink

Hyperlink can be added to pot objects.

Use argument `hyperlink` with function `pot`.

Here is an example:

```{r}
pot( 'Cats', hyperlink = 'http://en.wikipedia.org',
  textBoldItalic( color = '#428BCA', underline = TRUE ) ) 
```

## Footnotes

Footnotes can be added to pot objects.

Use objects of class `Footnote` to create a Footnote. Function `pot` has an 
argument `footnote` to specify which Footnote to use. If missing, no footnote 
will be added.

**Footnote are not supported within PowerPoint documents.**

### Word example

```{r}
library( ReporteRs )

footnote1 = Footnote( )
footnote1 = addParagraph( footnote1, pot('About this reference', textBold( ) ) )

mydoc = docx( )
mydoc = addParagraph(mydoc, pot('Latius iam disseminata licentia', footnote = footnote1 ) )

```


