---
title: "PowerPoint documents"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{PowerPoint documents}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
library(ReporteRs)
if(!require(filelink))
  as_link <- function(file, txt, label){invisible()}

options("ReporteRs-fontsize" = 11)
```


## Getting Started

### Create a pptx object

Use the function `pptx` to create an r object representing a PowerPoint document.
It takes two arguments: a title (appearing only in the document properties) and
a template file. If none is provided, the template file will be an empty document
located in the package directory.

When creating a pptx object, you use a template file. This file is copied in memory
and that copy becomes the document that will be completed with R outputs. Formats
and available slide layouts will be those available in the template file.

### Send R outputs into that object

You must **add a slide before you are able to send R outputs** into that
document (into that slide...). Use the function `addSlide` to add a 
slide into the **pptx** object. The number of available shapes 
that can be filled by R outputs is conditionned by the chosen 
layout. For example, layout 'Title and Content'
can receive only one R output, 'Two Content' can receive two R output. 

Next, you create the components of your slide. You then add text, tables, 
plots and other components into the slide that has just been added, available functions
are documented [below](#functions_list). 

*When adding an R output, if no position nor size is specified, the R output will be
placed with position and dimensions of the available shape defined in the template
layout. If you need a new layout, there are two options, create it in PowerPoint or 
[force size and position](#force_size_and_position).*

### Write the object to a file

Finally, using `writeDoc`, you write the object into a file
with the suffix *.pptx* on your file system.

### Example

Below you can follow this process with a commented R script:

```{r message=FALSE}
library( ReporteRs )

# Creation of doc, a pptx object (default template)
mydoc <- pptx( )

# check my layout names:
slide.layouts(mydoc)

mydoc <- addSlide( mydoc, "Two Content" )
# add into mydoc first 10 lines of iris
mydoc <- addTitle( mydoc, "First 10 lines of iris" )
mydoc <- addFlexTable( mydoc, vanilla.table(iris[1:10,] ) )

# add text into mydoc (and an empty line just before). Paragraph 
# properties will be those of the shape of the used layout.
mydoc <- addParagraph( mydoc, value = c("", "Hello World!") )

mydoc <- addSlide( mydoc, "Title and Content" )
# add a plot into mydoc 
mydoc <- addPlot( mydoc, function() barplot( 1:8, col = 1:8 ) )

filename <- tempfile(fileext = ".pptx") # the document to produce
# write mydoc 
writeDoc( mydoc, filename )
```

```{r echo=FALSE}
as_link(file = filename, txt = "It produces ", label = basename(filename) )
```


## Templates, layouts and styles

`pptx` will let you create documents based on a 'pptx' template file.

If none is provided, an empty document will be used (located into the package directory).

```{r eval=FALSE}
# use D:/docs/template/my_corporate_template.pptx as template
doc <- pptx(template = 'D:/docs/template/my_corporate_template.pptx')

# use default template
doc <- pptx()
```

`ReporteRs` use PowerPoint file as templates. A template is the original
PowerPoint document where all slides layouts, shapes (placeholders) and
styles come from:

* Slide layouts are those available in the template.
* Designs and formating properties of layouts and shapes
(placeholders in a layout) are defined within the template.
* Content of the PowerPoint template is not deleted, which lets
you to easily add content into an existing presentation
without usual annoying copy-paste operations.

**If you need other layouts than those that come with PowerPoint's
built-in Slide Master** of your template, you can add a layout and
customize it for your PowerPoint presentation. To add your own layout,
just follow these steps in PowerPoint:

1. Switch to the Slide Master view
2. Insert Layout.
3. Insert the placeholders you want to add to the new layout.
Move and or resize them regarding to the content it should contains.
4. You can apply any formatting to new placeholders (i.e. change background color, font, etc.)
5. Save and close the template
6. Create a `pptx` object with function pptx using the argument `template` specifying the template filename.


## Key functions and arguments

### Function addSlide

**If no slide has been added into the document object no content (tables,
plots, images, text) can be added.**

Funtion `addSlide`  let you add a slide in the pptx object. Argument `slide.layout` 
let you specify which layout to use to design the new slide.

```{r}
doc <- pptx()
doc <- addSlide( doc, slide.layout = 'Title and Content' )
doc <- addSlide( doc, slide.layout = 'Two Content' )
filename <- tempfile(fileext = ".pptx") # the document to produce
writeDoc( mydoc, filename )
```

```{r echo=FALSE}
as_link(file = filename, txt = "It produces ", label = basename(filename) )
```

### Function `slide.layouts`

Available layouts are those contained in the template. Slide layout concept is important in ReporteRs.

Content is conditioned by the chosen layout. When a slide is added
with layout 'Title and Content', this slide will only be able to
contain 2 contents: a title and a content theses are shapes.

Shapes (or placeholders) are where most of the formatting is done:
plot size is defined by shapes dimensions in the layout, default
fonts and paragraphs styles are those defined by the shape properties, etc.

`slide.layouts` returns slide layouts names. When adding a slide with `addSlide`
you may need that function to ckeck available names.

```{r}
doc <- pptx()
slide.layouts(doc)
```

**Check slide layout**

use optional argument `layout` with function `slide.layouts` to get
a plot of the design of your layout.

```{r}
doc <- pptx()
slide.layouts(doc, 'Comparison')
```

Here, the slide can receive the following shapes: a title, 4 contents
(plots, tables, paragraphs), a date and a slide number.

### Force size and position

Width, height and positions of the output (i.e. shape size and position)
are defined by the properties of the shape defined in the layout of the
PowerPoint template used.

**When adding content, ReporteRs read the properties of the next available
shape in the current slide.**

**You can explicitly specify shape position and dimensions for some specific
outputs: tables, text, plots and external images.**.

If arguments `offx`, `offy`, `width`, `height` are missing, position
and dimensions are defined by the width and height of the next available shape of the slide.

If these arguments are specified, they become position and dimensions of the
new shape. It let you also to add content on a slide if no more shape is available.

```{r}
library( ggplot2 )

mydoc <- pptx( title = "title" )

mydoc <- addSlide( mydoc, slide.layout = "Title and Content" )

myplot = qplot(Sepal.Length, Petal.Length, data = iris, 
  color = Species, size = Petal.Width, alpha = I(0.7) )

mydoc <- addPlot( doc = mydoc, fun = print, x = myplot, 
  offx = 1, offy = 1, width = 6, height = 5 )

mydoc <- addFlexTable( doc = mydoc, FlexTable( head( iris ) ),
  offx = 8, offy = 2, width = 4.5, height = 3 )

filename <- tempfile(fileext = ".pptx") # the document to produce
writeDoc( mydoc, file = filename )
```
```{r echo=FALSE}
as_link(file = filename, txt = "It produces ", label = basename(filename) )
```

## Specific functions

### Date

Function `addDate` lets you adding a date in the current slide.
The shape will be positioned with the date shape properties of the template document.

```{r}
doc <- pptx()
doc <- addSlide( doc, slide.layout = 'Title and Content' )
doc <- addDate( doc )
doc <- addSlide( doc, slide.layout = 'Title and Content' )
doc <- addDate( doc, 'Dummy date' )
filename <- tempfile(fileext = ".pptx") # the document to produce
writeDoc( mydoc, file = filename )
```
```{r echo=FALSE}
as_link(file = filename, txt = "It produces ", label = basename(filename) )
```

### Page number

Function `addPageNumber` lets you adding slide number in the current slide. The
shape will be positioned with the page number shape properties of the template document.

```{r}
doc <-pptx()
doc <-addSlide( doc, slide.layout = 'Title and Content' )
doc <-addPageNumber( doc )
doc <-addSlide( doc, slide.layout = 'Title and Content' )
doc <-addPageNumber( doc, 'Dummy text' )
filename <- tempfile(fileext = ".pptx") # the document to produce
writeDoc( mydoc, file = filename )
```
```{r echo=FALSE}
as_link(file = filename, txt = "It produces ", label = basename(filename) )
```

### Footer

Function `addFooter` lets you adding a comment in the footer of current slide.
The shape will be positioned with the footer shape properties of the template
document.

```{r}
doc <- pptx()
doc <- addSlide( doc, slide.layout = 'Title and Content' )
doc <- addFooter( doc, 'Dummy text' )
filename <- tempfile(fileext = ".pptx") # the document to produce
writeDoc( mydoc, file = filename )
```
```{r echo=FALSE}
as_link(file = filename, txt = "It produces ", label = basename(filename) )
```

### Subtitle

Function `addSubtitle` lets you adding a subtitle in the current slide. The
shape will be positioned with the subtitle shape properties of the template document.

Subtitle shape only exist in slide of type 'Title Slide'.

```{r}
doc <- pptx()
doc <- addSlide( doc, slide.layout = 'Title Slide' )
doc <- addTitle( doc, 'Presentation title' )
doc <- addSubtitle( doc , 'This document is generated with ReporteRs.')
filename <- tempfile(fileext = ".pptx") # the document to produce
writeDoc( mydoc, file = filename )
```
```{r echo=FALSE}
as_link(file = filename, txt = "It produces ", label = basename(filename) )
```


## Replace slide

To replace an existing slide by a new one, use `addSlide` with
optional argument bookmark. Bookmark is the index of the slide
that has to be replaced with a new empty one.

```{r eval=FALSE}
mydoc <-pptx( title = 'title', template = 'template_example.pptx' )

myplot = qplot(Sepal.Length, Petal.Length, data = iris, color = Species, size = Petal.Width, alpha = I(0.7))

mydoc <-addSlide( mydoc, slide.layout = 'Title and Content', bookmark = 3 )

mydoc <-addTitle( mydoc, 'my new graph')

mydoc <-addPlot( mydoc, print, x = myplot )

writeDoc( mydoc, 'pp_replacement.pptx' )

```


## Working with pipes

ReporteRs can be used with package `magrittr`, it makes code easier to read and more compact.

```{r message=FALSE}
library( magrittr )
filename <- tempfile(fileext = ".pptx") # the document to produce

pptx( title = "title" ) %>% 
  addSlide( slide.layout = "Title and Content" ) %>% 
  addFlexTable( vanilla.table(head(iris) ) ) %>% 
  writeDoc( file = filename )
filename <- tempfile(fileext = ".pptx") # the document to produce
writeDoc( mydoc, file = filename )
```
```{r echo=FALSE}
as_link(file = filename, txt = "Get the result ", label = "here" )
```


