---
title: "Word documents"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Word documents}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ReporteRs)
options("ReporteRs-fontsize" = 11)
```

## Getting Started

### Create a docx object
    
Use the function `docx` to create an r object representing a Word document. 
It takes two arguments: a title (appearing only in the Word document properties) 
and a **template file**. If none is provided, the template file will be an empty 
document located in the package directory.

When creating a docx object, you use a template file. This file is copied 
in memory and that copy becomes the document that will be completed with R outputs. 
Formats and available styles will be those available in the template file.

### Send R outputs into that object

Next, you create the components of your docx file. You create 
and format text, then insert it using the addParagraph function. 
You create and insert tables, perhaps merging cells, changing 
row and column colors or fonts. And, of course, you add plots. 
Tables, plots and paragraphs formats can be customised with 
dedicated functions.

### Write the object to a file

Finally, using `writeDoc`, you write the object into a file 
with the suffix `.docx` on your file system.

### Example

Below you can follow this process with a commented R script:

```{r}
library( ReporteRs )

# Creation of mydoc, a mydocx object
mydoc <- docx( )

# add into mydoc first 10 lines of iris
mydoc <- addFlexTable( mydoc, vanilla.table(iris[1:10,] ) )

# add a page break
mydoc <- addPageBreak( mydoc )

# add text with stylename "Normal" into mydoc 
mydoc <- addParagraph( mydoc, value = "Hello World!", stylename = "Normal" )

# add a plot into mydoc 
mydoc <- addPlot( mydoc, function() barplot( 1:8, col = 1:8 ) )

filename <- "base_example.docx" # the document to produce
# write the doc 
writeDoc( mydoc, file = filename)
```

```{r echo=FALSE}
office_web_viewer( url = paste0( "http://davidgohel.github.io/ReporteRs/articles/", basename(filename) ) )
```


## Templates

You may need to generate docx documents in a particular corporate template 
(with specific fonts, colour schemes, logos, etc.). The function `docx` 
has an optional argument **template**, it lets you create documents based 
on existing Word documents. 

If you don't direct ReporteRs to a template file, an empty document 
will be used (located into the package directory) and you will have a document 
with its Word styles.

**It loads in memory your template word document to add content to**. 
R outputs will be inserted at the end of that copy (except if you are using bookmarks).
The approach is to use a Word document that contains the layout and main styles of 
your final document. You can edit an existing file in Word, delete the whole content 
or not, modify paragraph styles, headers, etc., and use it as a template. New 
reports will have the same layout and styles than the template.


```{r eval=FALSE}
# use D:/docs/template/my_corporate_template.docx as template
doc <- docx(template = 'D:/docs/template/my_corporate_template.docx')

# use default template
doc <- docx()
```

If template content should be deleted, use the argument  `empty_template=TRUE` when calling `docx` function:

```{r eval=FALSE}
doc <- docx(template = 'D:/docs/template/my_corporate_template.docx', empty_template = TRUE)
```

## Styles 

Available styles will be paragraph styles of the base document 
(e.g. Normal, Title1, etc.). Names of the returned character vector 
are internal Word labels associated with styles names.

```{r}
doc <- docx()
styles( doc )
```

*When using a template, styles names of the template should only 
contain letters (from a to z) and numbers. For example, in France 
the default style named `Legend` becomes `L?gende` 
and is returned as `Lgende` by the styles R function. The workaround 
is to create a new style based on `L?gende` and to name it `Legende`. 
It will be then a valid name to use with ReporteRs.*

**styles** are used by several function, adding text with `addParagraph`, 
adding titles with `addTitle`, adding specific TOC with `addTOC`.

### Paragraphs and styles

Function `addParagraph` has an optional argument `stylename` whose value must 
be one of the names returned by the function `styles`.


```{r}
doc <- docx( )

# display all available styles with their names
doc <- addParagraph( doc, value = "Hello", stylename = "Normal")

filename <- "styles.docx" # the document to produce
writeDoc( doc, file = filename)
```

```{r echo=FALSE}
office_web_viewer( url = paste0( "http://davidgohel.github.io/ReporteRs/articles/", basename(filename) ) )
```

## Specific functions

### Table of contents

Insert a table of contents with function `addTOC`.

```{r eval=FALSE}
doc <- addTOC(doc)
```

> When added, a TOC will make Word to pop-up a message-box asking you
if you want to update TOC entries. **This is not an error**, you should
click 'Yes' to update TOC entries.

#### Customized table of contents

Add a customized table of contents with the argument `stylename`.
If used, a table of contents will be created containing entries
formatted with specified styles.

```{r eval=FALSE}
# add a table of all paragraphs with style 'stylename'
doc <- addTOC( doc, stylename = 'stylename' )
```

**Example: Add a list of figures and a list of tables**

In the following example, style `figurereference` is used as style of paragraphs
following graphics, as a **caption** for plots. Style `tablereference` is used as style
of paragraphs following table, as a **caption** for tables.

These styles are available in the template file *templates/template_toc.docx*.

We will:

1. add a table of content
2. add a table of figures
3. add a table of tables

```{r message=FALSE}
library( ReporteRs )
library( ggplot2 )
library(magrittr)

myplot1 <- qplot(Sepal.Length, Petal.Length, 
  data = iris, color = Species, 
  size = Petal.Width, alpha = I(0.7))
myplot2 <- qplot(mpg, wt, data = mtcars, colour = cyl )

# Create a new document
mydoc <- docx( template = "template_toc.docx" )

# display available styles
styles( mydoc )

mydoc <- mydoc %>% addTitle( "Table of contents", level =  1 ) %>% 
  addTOC( ) %>% # add a table of content
  addPageBreak() # add a page break

mydoc <- mydoc %>% addTitle( "List of graphics", level =  1 ) %>% 
  addTOC( stylename = "figurereference" ) %>% 
  addTitle( "List of tables", level =  1 ) %>% 
  addTOC( stylename = "tablereference" )

mydoc <- addPageBreak( mydoc )

mydoc <- mydoc %>% addTitle( "iris outputs", level =  1 ) %>% 
  addTitle( "Plot", level =  2 ) %>% 
  addPlot(function( ) print( myplot1 ) ) %>% 
  addParagraph( value = "iris plot", 
    stylename = "figurereference") %>%  # Add a legend below the plot
  addTitle( "Table", level =  2 ) %>% 
  addFlexTable( vanilla.table( head( iris ) ) ) %>%
  addParagraph( value = "iris table", 
    stylename = "tablereference") # Add a legend below the table

mydoc <- mydoc %>% addTitle( "mtcars outputs", level =  1 ) %>% 
  addTitle( "Plot", level =  2 ) %>% 
  addPlot(function( ) print( myplot2 ) ) %>% 
  addParagraph( value = "mtcars plot", 
    stylename = "figurereference") %>%  # Add a legend below the plot
  addTitle( "Table", level =  2 ) %>% 
  addFlexTable( vanilla.table( head( mtcars ) ) ) %>%
  addParagraph( value = "mtcars table", 
    stylename = "tablereference") # Add a legend below the table

filename <- "toc.docx" # the document to produce
writeDoc( doc, file = filename)
```

```{r echo=FALSE}
office_web_viewer( url = paste0( "http://davidgohel.github.io/ReporteRs/articles/", basename(filename) ) )
```

### Page breaks

Insert a page break with function addPageBreak

    
```{r eval=FALSE}
doc <- addPageBreak(doc)
```

### Section

Function `addSection` lets you add a section. 
With a section, you can change document orientation and 
split new content along 2 or more columns.

The function requires you to **add a section before 
and after** the item(s) that you want to be on a 
landscape and/or multicolumns mode page. 

```{r message=FALSE}
library(magrittr)

doc <- docx()

doc <- addSection( doc, landscape = TRUE) %>% 
  addFlexTable( vanilla.table(mtcars)) %>% 
  addSection()

doc <- addSection( doc, ncol = 2, landscape = TRUE) %>% 
  addParagraph( "My text on column 1", stylename = "Normal" ) %>% 
  addColumnBreak( ) %>% 
  addParagraph("Hi on the right", stylename = "Normal" ) %>% 
  addSection( )

filename <- "section.docx" # the document to produce
writeDoc( doc, file = filename)
```

```{r echo=FALSE}
office_web_viewer( url = paste0( "http://davidgohel.github.io/ReporteRs/articles/", basename(filename) ) )
```


